\speculake: Hiding Malware in Speculative Execution

### 2018-08-02 paper/abstract.tex
preliminary discussion update

We introduce \emph{\speculake}\footnote{Executable Spectre}, which compiles
arbitrary malicious code into a seemingly-benign payload binary. When a separate
trigger program runs on the same machine, it mistrains the CPU's branch
predictor, causing the payload program to \emph{speculatively} execute code, and
communicating its results back to the rest of the payload program, changing its
behavior.

>>>>

We introduce \emph{\speculake}\footnote{Executable Spectre}, which compiles
arbitrary malicious code into a seemingly-benign payload binary. When a separate
trigger program runs on the same machine, it mistrains the CPU's branch
predictor, causing the payload program to \emph{speculatively} execute code, and
communicating its results back to the rest of the payload program, changes its
behavior.

### 2018-08-02 paper/architecture.tex
preliminary discussion update

\speculake malware is comprised of two independent programs: a payload program,
and a trigger program. Both are installed on the victim's computer (e.g. via
trojan, remote exploit, or phishing), and must run simultaenously on the same
physical CPU. We note that the constraint of running programs on the same CPU is
not a signifcant burden: \texttt{taskset} can be used to limit a process to a
core, or if not available, the attacker can simply run multiple copies of the
trigger or payload program to coerce the OS scheduler to assign a
payload/trigger pair to the same core.

@@

We further assume that the malware has a specific trigger that the analyst is
not privy to. In our examples, this trigger is often behavior exhibited by some
other (potentially benign) process running on the same system as the malware.
As the adversary is able to control when such a trigger is deployed (potentially
remotely), the analyst will not be able to observe or force this trigger to
happen at will. We emphasize that while this may also be true for existing
trigger-based malware, analysts can often reverse engineer the trigger out of
the malware, for example by observing control flow within the malware and using
adaptive fuzzers~\cite{afl} to generate inputs that explore other execution
paths of the binary. In contrast, \speculake malware's trigger depends on
behavior in seemingly unrelated binaries, influencing malware behavior through
channels invisible to the analyst. As with typical malware, the analyst may
attempt to reverse engineer the trigger to revela the malware's behavior, but we
will show (in Section~\ref{subsec:nested-spec}) how to defeat this type of
analysis.

@@

We performed several experiments to determine how much computation can be
performed
speculatively, as well as what components are responsible for the limit.
We report results from our experiments on an Intel~Xeon-1270,
though we note we found similar results across other Intel processor generations,
including an i5-7200U, % jack
an i5-4300U, % ewust laptop
and an i5-4590. % ewust desktop
% For simplicity, we report results from only the Xeon-1270.

@@

\subsubsection{Hyperthreading}

@@

>>>>

\speculake malware is comprised of two independent programs: a payload program,
and a trigger program. Both are installed on the victim's computer (e.g. via
trojan, remote exploit, or phishing), and must run simultaenously on the same
physical CPU. We note that the constraint of running programs on the same CPU is
not a signifcant burden: \texttt{taskset} can be used to limit a process to a
core, or if not available, the attacker can simply run multiple copies of the
trigger or payload program to coerce the OS scheduler to assign a
payload/trigger pair to the same core. We note that this is not a strict requirement
for the attacker as most OS schedulers will use round robin algrithms to assign 
tasks to processors -- eventually the trigger and payload will execute co-resident
on a single core.

@@

We further assume that the malware has a specific trigger that the analyst is
not privy to. In our examples, this trigger is often behavior exhibited by some
other (potentially benign) process running on the same system as the malware.
As the adversary is able to control when such a trigger is deployed (potentially
remotely), the analyst will not be able to observe or force this trigger to
happen at will. We emphasize that while this may also be true for existing
trigger-based malware, analysts can often reverse engineer the trigger out of
the malware, for example by observing control flow within the malware and using
adaptive fuzzers~\cite{afl} to generate inputs that explore other execution
paths of the binary. In contrast, \speculake malware's trigger depends on
behavior in seemingly unrelated binaries, influencing malware behavior through
channels invisible to the analyst. As with typical malware, the analyst may
attempt to reverse engineer the trigger to reveal the malware's behavior, but we
will show (in Section~\ref{subsec:nested-spec}) how to defeat this type of
analysis.

@@

We performed several experiments to determine how much computation can be
performed
speculatively, as well as what components are responsible for the limit.
We report results from our experiments on an Intel~Xeon-1270,
though we note we found similar results across other Intel processor generations,
including an i5-7200U, % jack
an i5-4300U, % ewust laptop
an i5-4590, % ewust desktop
and an Epyc 7251.
% For simplicity, we report results from only the Xeon-1270.

@@

\subsubsection{Hyperthreading}
\label{sssec:hyperthread}

@@

We note that Single Thread Indirect Branch Predictors (STIBP) have been implemented
in most environments to prevent cross thread branch predictor interference.
While this does remove the \speculake trigger's influence in scenarios where a
process runs on an isolated cpu, in most modern environments
tasks are scheduled to all available processors. This means that the trigger and
payload will be coresident eventually, allowing progress to continue.
